---
title: "Ranking ordered data tutorial"
author: "Jose Nieves"
date: "March 1, 2016"
output: html_document
---

#How to return a specific ranking after ordering

Working on cities.csv, we want to know what's the __second biggest city in our UK database__. We are not interested in an ordered data frame, we just want to know the name of the city.

```{r}
data <- read.csv("cities.csv")
data
class(data)
```

#### __argument decreasing = TRUE__ inverts the direction of the order. Numbers from biggest to smallest and characters from Z to A. This is helpful when we consider rank1 the biggest city.
#### __as.character ( )__ will return the vector with the name of the city. If we just return orderdata[rank,1] we get a factor instead.

```{r}
find_city_rank <- function(data,column,rank){
    orderdata <- data[order(decreasing = TRUE,data[,column]),]
    return(as.character(orderdata[rank,1]))
}
```

##Examples:

```{r}
find_city_rank (data,3,1)
find_city_rank (data,3,2)
find_city_rank (data,4,2)
```

Now let's consider the case in which we __don't know the length of the csv file and we just want to get the last city in the ranking.__

####__nrow( )__ returns the number of rows of a data frame. We use this function to determine the index of the last item.

```{r}
find_last_city <- function(data,column){
    orderdata <- data[order(decreasing = TRUE,data[,column]),]
    return(as.character(orderdata[nrow(orderdata),1]))
}
```

##Examples:
```{r}
find_last_city (data,3)
find_last_city (data,1)
```
####__Note:__ that in the last example Beijing it's the last in alphabetical ranking. This is because decreasing = TRUE.

