---
title: "Order data frames with NA's tutorial"
author: "Jose Nieves"
date: "March 1, 2016"
output: html_document
---

# How to manage NA when ordering data frames

Let's read citiesNA.csv file

```{r}
data <- read.csv("citiesNA.csv")
data
class(data)
```    

In this case we have some character data in the areakm2 and populationk that we want to consider as NA.

Below is the function to order the data frame by column, that considers only complete rows with no NA values.

This means that we will exclude NYC and London form our data frame.

##Subsetting data by column, we get a factor:
```{r}
class(data[,2])
```
## One way to extract a vector from the factor is by subsetting it by its levels.
###__levels(data[,2])__ returns a vector of the levels:
```{r}
levels(data[,2])
```

###__levels(data[,2])[data[,2]]__ returns a vector with the content of [data[,2]]

```{r}
levels(data[,2])[data[,2]]
```

###__data[,2]__ would return a factor, that for our purposes is harder to handle

```{r}
data[,2]
```    

__SuppressWarnings()__ stops the warning alerts from R. When we coerce a mixed list of numeric and character into a numeric vector, text becomes automatically NA, but it's a forced coercion and R sends a warning. This is the case of our column 3 and 4, where "Unknown" becomes NA.

__complete.cases()__ returns the indexes of the rows that don't have any NA. By subsetting the matrix by these indexes we get a data frame with only complete cases.
```{r}
sort_by_column_NA <- function(data,column) {
        for (i in 3:4){
                 data[,i] <- suppressWarnings(as.numeric(levels(data[,i])[data[,i]]))
        }
        orderdata <- data[order(data[,column]),]
        orderdata <- orderdata[complete.cases(orderdata),] 
        return(orderdata)
}
```

##Examples:

```{r}
sort_by_column_NA(data,3)
```
       
```{r}
sort_by_column_NA(data,1)
```
       