---
title: "Order data frames tutorial"
author: "Jose Nieves"
date: "February 29, 2016"
output: html_document
---

# How to order a data frame by the values in its columns

Let's read first the cities.csv file
```{r}
data <- read.csv("cities.csv")
data
class(data)
```

As we can see data is a data.frame with 6 rows and 4 columns, both character and numeric.

Below is the function to order the data frame by column:

orderdata: output data.frame with the ordered rows
__order()__: sort by default in decreasing order the values.

In case of numbers from the smallest to the biggest, in case of characters from A to Z. 

####This function returns a vector of indexes with the ordered rows. 
```{r}
order(data[,1])
```

####__data[order(),]__ subsets the data frame using the indexes above

```{r}
sort_by_column <- function (data, column){
        orderdata <- data[order(data[,column]),]
        return(orderdata)
}
```

##Examples:

```{r}
sort_by_column (data,1)
sort_by_column (data,3)
```

####In case of __tie__, we might consider to give a second attribute to order() to give a second criteria

```{r}
sort_by_columns <- function (data, col1, col2){
        orderdata <- data[order(data[,col1],data[,col2]),]
        return(orderdata)
}
```

##Examples:

```{r}
sort_by_columns (data,2,3)
sort_by_columns (data,2,1)
```	

