---
title: "Subsetting tutorial"
author: "Jose Nieves"
date: "March 1, 2016"
output: html_document
---

#How to order only a specific subset of a data frame
Let's read cities.csv.
```{r}
data <- read.csv("cities.csv")
data
class(data)
```
If our purpose is to see what is the __largest city in China between Shanghai and Beijing__, we might be not interested in the other cities'areas. 

To do so we subset our initial data frame, and then we order it considering the criteria we prefer. It doesn't make much sense with 2 items, but when we are analysing thousands of items it's far more useful.

### __grep( )__ finds the character vector (e.g. "China") in the data$countries factor, and returns a vector of indexes.
```{r}
data$countries
  
grep("China",data$countries)

## We then subset the main data frame, data, by these indexes

data [grep("China",data$countries),]
```

```{r}
sort_country <- function (data, country, column){
  countrydata <- data [grep(country,data$countries),]
  orderdata <- countrydata[order(countrydata[,column]),]
  return (orderdata)
}
```

##Examples:

```{r}
sort_country(data, "USA", 4)
```
  
